---
title: "One iteration, ash posterior mean"
author: "Dongyue Xie"
date: "May 8, 2018"
output: html_document
---
```{r}
library(smashr)
library(ashr)
```

```{r}
#' smash generaliation function, expansion around ash posterior mean


#' @param x: a vector of observations
#' @param sigma: standard deviations, scalar.
#' @param family: choice of wavelet basis to be used, as in wavethresh.
#' @param niter: number of iterations for IRLS
#' @param tol: tolerance of the criterion to stop the iterations
#' @param ashp: whether expand around ash posterior mean 
#' @param robust: whether set the highest resolution wavelet coeffs to 0
#' @param verbose: whether print out the number of iterations to converge.

smash.gen=function(x,sigma,family='DaubExPhase',
                   ashp=TRUE,verbose=FALSE, robust=FALSE,
                   niter=30,tol=1e-2){
  mu=c()
  s=c()
  y=c()
  munorm=c()
  #apply ash to poisson data?
  if(ashp){
    pmean=ash(rep(0,length(x)),1,lik=lik_pois(x))$result$PosteriorMean
    mu=rbind(mu,pmean)
    s=rbind(s,1/mu)
    y0=log(mu)+(x-mu)/mu
  }else{
    mu=rbind(mu,rep(mean(x),length(x)))
    s=rbind(s,1/mu)
    y0=log(mean(x))+(x-mean(x))/mean(x)
  }
  #set wavelet coeffs to 0?
  if(robust){
    wds=wd(y0,family = family,filter.number = filter.number)
    wtd=threshold(wds, levels = wds$nlevels-1,  policy="manual",value = Inf) 
    y=rbind(y,wr(wtd))
  }else{
    y=rbind(y,y0)
  }
  for(i in 1:niter){
    vars=ifelse(s[i,]<0,1e-8,s[i,])
    mu.hat=smash.gaus(y[i,],sigma=sqrt(vars))#mu.hat is \mu_t+E(u_t|y)
    mu=rbind(mu,mu.hat)
    munorm[i]=norm(mu.hat-mu[i,],'2')
    if(munorm[i]<tol){
      if(verbose){
        message(sprintf('Converge after %i iterations',i))
      }
      break
    }
    #update m and s_t
    mt=exp(mu.hat)
    s=rbind(s,1/mt)
    y=rbind(y,log(mt)+(x-mt)/mt)
    
    
  }
  mu.hat=smash.gaus(y[i,],sigma = sqrt(sigma^2+ifelse(s[i,]<0,1e-8,s[i,])))
  return(list(mu.hat=mu.hat,mu=mu,s=s,y=y,munorm=munorm))
}

```

```{r}
#' Simulation study comparing smash and smashgen

simu_study=function(m,sigma,seed=1234,
                    niter=1,family='DaubExPhase',ashp=TRUE,verbose=FALSE,robust=FALSE,
                    tol=1e-2,reflect=FALSE){
  set.seed(seed)
  lamda=exp(m+rnorm(length(m),0,sigma))
  x=rpois(length(m),lamda)
  #fit data
  smash.out=smash.poiss(x,reflect=reflect)
  smash.gen.out=smash.gen(x,sigma=sigma,niter=niter,family = family,tol=tol,ashp=ashp,verbose=verbose)
  return(list(smash.out=smash.out,smash.gen.out=exp(smash.gen.out$mu.hat),smash.gen.est=smash.gen.out,x=x,loglik=smash.gen.out$loglik))
}

```


# Simulation 1: Constant trend Poisson nugget

## $\sigma=0.01$

```{r}
library(smashr)
m=rep(3,256)
simu.out=simu_study(m,0.01)
simu.out.conv=simu_study(m,0.01,niter = 30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)

lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))

```

## $\sigma=0.1$

```{r}

simu.out=simu_study(m,0.1)
simu.out.conv=simu_study(m,0.1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))


```


## $\sigma=0.5$

```{r}
simu.out=simu_study(m,0.5)
simu.out.conv=simu_study(m,0.5,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)

lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))


```

## $\sigma=1$

```{r}

simu.out=simu_study(m,1)
simu.out.conv=simu_study(m,1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)

lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))


```

# Simulation 2: Step trend 

## $\sigma=0.01$

```{r}
m=c(rep(3,128), rep(5, 128), rep(6, 128), rep(3, 128))
simu.out=simu_study(m,0.01)
simu.out.conv=simu_study(m,0.01,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.1$

```{r}
simu.out=simu_study(m,0.1)
simu.out.conv=simu_study(m,0.1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.5$

```{r}
simu.out=simu_study(m,0.5)
simu.out.conv=simu_study(m,0.5,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=1$

```{r}

simu.out=simu_study(m,1)
simu.out.conv=simu_study(m,1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

# Simulation 3: Oscillating Poisson nugget


## $\sigma=0.01$
```{r}
m=c()
for(k in 1:8){
    m=c(m, rep(1,15), rep(5, 15))
}
m=c(m,rep(1,16))
simu.out=simu_study(m,0.01)
simu.out.conv=simu_study(m,0.01,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
# legend("topleft", # places a legend at the appropriate place
#        c("truth","smash-gen-1iter","converged"), # puts text in the legend
#        lty=c(1,1,1), # gives the legend appropriate symbols (lines)
#        lwd=c(1,1,1),
#        cex = 1,
#        col=c("black","red", "blue"))
```

## $\sigma=0.1$
```{r}
simu.out=simu_study(m,0.1)
simu.out.conv=simu_study(m,0.1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.5$
```{r}
simu.out=simu_study(m,0.5)
simu.out.conv=simu_study(m,0.5,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=1$
```{r}
simu.out=simu_study(m,1)
simu.out.conv=simu_study(m,1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

# Simulation 4: Bumps

## $\sigma=0.01$
```{r}
m=seq(0,1,length.out = 256)
h = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
w = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005,0.008,0.005)
t=c(.1,.13,.15,.23,.25,.4,.44,.65,.76,.78,.81)
f = c()
for(i in 1:length(m)){
  f[i]=sum(h*(1+((m[i]-t)/w)^4)^(-1))
}
m=f
simu.out=simu_study(m,0.01)
simu.out.conv=simu_study(m,0.01,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.1$
```{r}
simu.out=simu_study(m,0.1)
simu.out.conv=simu_study(m,0.1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.5$
```{r}
simu.out=simu_study(m,0.5)
simu.out.conv=simu_study(m,0.5,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=1$
```{r}
simu.out=simu_study(m,1)
simu.out.conv=simu_study(m,1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

# Simulation 5: Sin curve Poisson nugget

## $\sigma=0.01$
```{r}
m = seq(-pi,pi,length.out = 256)
m = 2*(sin(m)+1)
simu.out=simu_study(m,0.01)
simu.out.conv=simu_study(m,0.01,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.1$
```{r}
simu.out=simu_study(m,0.1)
simu.out.conv=simu_study(m,0.1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=0.5$
```{r}
simu.out=simu_study(m,0.5)
simu.out.conv=simu_study(m,0.5,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```

## $\sigma=1$
```{r}
simu.out=simu_study(m,1)
simu.out.conv=simu_study(m,1,niter=30)
#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out.conv$smash.gen.out, col = "blue", lwd = 2)
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", # places a legend at the appropriate place
       c("truth","smash-gen-1iter","converged"), # puts text in the legend
       lty=c(1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1,1),
       cex = 1,
       col=c("black","red", "blue"))
```


# Summary

By applying ash to Poisson data first and expanding around the estimated posterior mean, the performance of 1 iteration algorithm is greatly improved - actually from the above plots, it is almost the same as converged version.
