---
title: "Binomial wavelet smoothing"
author: "Dongyue Xie"
date: "May 12, 2018"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Algorithm

Let $X_t$ be a Binomial observation with number of trials $n_t$, $t=1,2,\dots,T$. 

1. Apply `ash` to binomial data $(X_t, n_t)$ and obtain the posterior mean $\hat p=E(p|X,n)$, $\hat p\in R^T$.
2. Let $s_t=\frac{1}{n_t\hat p_t(1-\hat p_t)}$ and $Y_t=\log\frac{\hat p_t}{1-\hat p_t}+\frac{X_t/n_t-\hat p_t}{n_t\hat p_t(1-\hat p_t)}$.
3. Apply `smash.gaus` to $Y_t$ with standard deviation $\sqrt(\sigma^2+s_t^2)$ if $\sigma^2$ is known. If it's unknown, then either $\sigma^2$ or $\sigma^2+s_t^2$ should be estimated first. The output of `smash.gaus` is denoted as $\mu_t$.
4. Export $\frac{\exp(\mu_t)}{1+\exp(\mu_t)}$.

Methods for estimating $\sigma^2$ or $\sigma^2+s_t^2$ are discussed [here](sigma.html).

```{r}
library(smashrgen)
library(ggplot2)
simu_study=function(p,sigma,ntri,nsimu=100,seed=12345,
                    niter=1,family='DaubExPhase',ashp=TRUE,verbose=FALSE,robust=FALSE,
                    tol=1e-2){
  set.seed(seed)
  #smash.err=c()
  smashgen.err=c()
  smashgen.smashu.err=c()
  #smashgen.mle.err=c()
  n=length(p)
  true.p=exp(p)/(1+exp(p))
  for(k in 1:nsimu){
    ng=rnorm(n,0,sigma)
    m=exp(p+ng)
    q=m/(1+m)
    x=rbinom(n,ntri,q)
    #fit data
    #smash.out=smash.poiss(x)
    smashgen.out=smash_gen(x,dist_family = 'binomial',sigma = sigma,ntri=ntri)
    smashu.out=smash_gen(x,dist_family = 'binomial',y_var_est = 'smashu',ntri=ntri)
    #mle.out=smash_gen(x,dist_family = 'poisson',y_var_est = 'mle')
    #smash.err[k]=mse(exp(m),smash.out)
    smashgen.err[k]=mse(true.p,smashgen.out)
    smashgen.smashu.err[k]=mse(true.p,smashu.out)
    #smashgen.mle.err[k]=mse(exp(m),mle.out)
  }
  return(list(est=list(smashgen.out=smashgen.out,smashu.out=smashu.out,x=x),err=data.frame(smashgen=smashgen.err, smashgen.smashu=smashgen.smashu.err)))
}
```

# Constant trend

## $\sigma=0.1$

```{r const}
p=rep(0,128)
set.seed(111)
ntri=rpois(128,30)
result=simu_study(p,0.1,ntri)
par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
legend("topright", # places a legend at the appropriate place
       c("truth","smashgen"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)
legend("topright", # places a legend at the appropriate place
       c("truth","smashgen"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("black","blue"))

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$

```{r const1}
result=simu_study(p,1,ntri)
par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$, small prob

```{r}
p=rep(-3,128)
result=simu_study(p,1,ntri)
par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

# Step trend

## $\sigma=0.1$
```{r}
p=c(rep(-2,128), rep(0, 128), rep(2, 128), rep(-2, 128))
set.seed(111)
ntri=rpois(512,30)
result=simu_study(p,0.1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$
```{r}
p=c(rep(-2,128), rep(0, 128), rep(2, 128), rep(-2, 128))
set.seed(111)
ntri=rpois(512,30)
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$, more variant ntri

```{r}
p=c(rep(-2,128), rep(0, 128), rep(2, 128), rep(-2, 128))
set.seed(111)
ntri=rpois(512,50)-20
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

# Bumps

## $\sigma=0.1$

```{r bumps}
m=seq(0,1,length.out = 256)
h = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
w = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005,0.008,0.005)
t=c(.1,.13,.15,.23,.25,.4,.44,.65,.76,.78,.81)
f = c()
for(i in 1:length(m)){
  f[i]=sum(h*(1+((m[i]-t)/w)^4)^(-1))
}
p=f-3

set.seed(111)
ntri=rpois(256,30)
result=simu_study(p,0.1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$

```{r bump}
set.seed(111)
ntri=rpois(256,30)
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$, more variant ntri

```{r}
set.seed(111)
ntri=rpois(256,50)-20
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$, smaller ntri

```{r}
set.seed(111)
ntri=rpois(256,20)-9
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```


# Spike mean

## $\sigma=0.1$

```{r spike}
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 256
t = 1:n/n
p = spike.f(t)*2-2

set.seed(111)
ntri=rpois(256,20)-9
result=simu_study(p,0.1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```


## $\sigma=1$

```{r spike1}
set.seed(111)
ntri=rpois(256,20)-9
result=simu_study(p,1,ntri)

par(mfrow=c(1,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: known variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x/ntri,col='gray80',ylab='',main='smashgen: unknown variance')
lines(exp(p)/(1+exp(p)),col=1)
lines(result$est$smashu.out,col=4)

mean(result$err$smashgen)
mean(result$err$smashgen.smashu)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```
