---
title: "Chip-exo data smoothing(cut)"
author: "Dongyue Xie"
date: "June 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

The lenght of sequence here is 425, which is not a power of 2 hence causing problem when using smashgen. In this real data analysis, we cut the data both to the left and right so that it has length of 256 then apply smashgen to the cuted data.

# Setup
```{r}
library(smashrgen)
library(dashr)

extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}

```

# EncodeBroadHistoneHelas - Rep 1

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
#y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(chipexo1$counts[1:256])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(1:256,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
#y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(chipexo1$counts[170:425])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(170:425,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# EncodeOpenChromChipHelas - Rep 1

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)

smashgen.out=smash_gen_lite(chipexo1$counts[45:300])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(45:300,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
#y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(chipexo1$counts[100:355])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(100:355,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# EncodeOpenChromChipHelas - Rep 2

##Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt")
dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
#y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(chipexo1$counts[1:256])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(1:256,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_reverse_counts.txt")
dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
smashgen.out=smash_gen_lite(chipexo1$counts[100:355])

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(100:355,smashgen.out,col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```
