---
title: "CTCF Chip-seq data smoothing"
author: "Dongyue Xie"
date: "June 3, 2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

The lenght of sequence here is 425, which is not a power of 2 hence causing problem when using smashgen. In this real data analysis, we reflect the data both to the left and right so that it has length of 1024 then apply smashgen to the augmented data.

# Setup

```{r}
library(smashrgen)
library(dashr)

extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}

```

# CTCF raw - Rep 1

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep1_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep1_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# CTCF raw - Rep 2

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep2_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep2_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# CTCF raw - Rep 3

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep3_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_raw_rep3_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# CTCF - Rep 1

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep1_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep1_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# CTCF - Rep 2

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep2_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep2_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

# CTCF - Rep 3

## Forward

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep3_Forward_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```

## Reverse

```{r}
chipexo1 <- extract_counts_CTCF("D:/smashgen/data/chipseq_examples/example_CTCF_rep3_Reverse_counts.txt")

dash.out=dash_smooth(chipexo1$counts, dash_control = list(Inf_weight = 1), progressbar = FALSE)
smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y)

plot(chipexo1$counts, col = "gray80", type = "h", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = "blue", lwd = 1)
lines(dash.out$estimate, col = "red", lwd = 2)
lines(smashgen.out[301:725],col=3,lwd=3)

legend("topright", # places a legend at the appropriate place
       c("truth","dash-m", "smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(1,1,1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2,1,3),
       cex = 0.5,
       col=c("gray80","red", "blue",3))
```
