---
title: "Missing data"
author: "Dongyue Xie"
date: "May 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

Treat unevenly spaced data as partly missing data. Set corresponding $s_t$ to $10^6$. The missing data are set to 0.

# Gaussian 

$Y_t=\mu_t+N(0,s_t^2)$

```{r}
library(smashrgen)
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 256
t = 1:n/n
mu = spike.f(t)+1
rsnr=2
var2 = (1e-04 + 4 * (exp(-550 * (t - 0.2)^2) + exp(-200 * (t - 0.5)^2) + exp(-950 * (t - 0.8)^2)))/1.35
sigma.ini = sqrt(var2)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu)/rsnr^2
set.seed(12345)
y=mu+rnorm(n,0,sigma.t)

# No missing data fit
fit=smash.gaus(y,sigma=sigma.t)
plot(y,type='p',col='gray80',main='No missing data')
lines(mu)
lines(fit,type='l',col=4)

# missing data 1%
idx=sample(1:n,n*0.01)

idx

y.miss=y
y.miss[idx]=0

sigma.miss=sigma.t
sigma.miss[idx]=10^6

fit=smash.gaus(y.miss,sigma=sigma.miss)
plot(y.miss,type='p',col='gray80',main='Missing 1% data')
lines(fit,col=4)
lines(idx,fit[idx],type='p',pch=16,col=2)
lines(mu,type='l')
legend('topright',c('smash.gaus','Missing points','True mu'),col=c(4,2,1),pch=c(NA,16,NA),lty=c(1,NA,1))

# # pretend variance are unknown
# fit2=smash.gaus(y.miss)
# plot(fit2,type='l',col=4,ylim=c(-0.5,3),main='Missing 1% data, variance unknown')
# lines(mu)

# missing data 5%
idx=sample(1:n,n*0.05)

sort(idx)

y.miss=y
y.miss[idx]=0

sigma.miss=sigma.t
sigma.miss[idx]=10^6

fit=smash.gaus(y.miss,sigma=sigma.miss)
plot(y.miss,type='p',col='gray80',main='Missing 5% data')
lines(fit,col=4)
lines(idx,fit[idx],type='p',pch=16,col=2)
lines(mu,type='l')
legend('topright',c('smash.gaus','Missing points','True mu'),col=c(4,2,1),pch=c(NA,16,NA),lty=c(1,NA,1))

# # pretend variance are unknown
# fit2=smash.gaus(y.miss)
# plot(fit2,type='l',col=4,ylim=c(-0.5,3),main='Missing 5% data, variance unknown')
# lines(mu)

# missing data 10%
idx=sample(1:n,n*0.1)

sort(idx)

y.miss=y
y.miss[idx]=0

sigma.miss=sigma.t
sigma.miss[idx]=10^6

fit=smash.gaus(y.miss,sigma=sigma.miss)
plot(y.miss,type='p',col='gray80',main='Missing 10% data')
lines(fit,col=4)
lines(idx,fit[idx],type='p',pch=16,col=2)
lines(mu,type='l')
legend('topright',c('smash.gaus','Missing points','True mu'),col=c(4,2,1),pch=c(NA,16,NA),lty=c(1,NA,1))

# missing data 20%
idx=sample(1:n,n*0.2)

sort(idx)

y.miss=y
y.miss[idx]=0

sigma.miss=sigma.t
sigma.miss[idx]=10^6

fit=smash.gaus(y.miss,sigma=sigma.miss)
plot(y.miss,type='p',col='gray80',main='Missing 20% data')
lines(fit,col=4)
lines(idx,fit[idx],type='p',pch=16,col=2)
lines(mu,type='l')
legend('topright',c('smash.gaus','Missing points','True mu'),col=c(4,2,1),pch=c(NA,16,NA),lty=c(1,NA,1))
```


# Why it's not working?

$y=\mu+\epsilon$

$Wy=W\mu+W\epsilon \Rightarrow\tilde y=\tilde\mu+N(0,diag(\tilde\sigma^2))$

We are shrinking the wavelet coefficients($\tilde\mu$) not $y$! If we set the missing data's $s_t$ very large, then all the variance of corresponding wavelet coefficients that involve missing data are very large. Thus, a large number of wavelet coefficients(essentially differences) are shrinked to zero, including those should not be done so.

An simple example:
```{r}
mu=c(1,1,1,1,4,4,4,4)
y=mu+rnorm(8,0,0.8)
sigma=rep(0.8,8)

w=t(GenW(filter.number = 1,family='DaubExPhase'))

y.miss=y
y.miss[6]=0
sigma.miss=sigma
sigma.miss[6]=1e6
y.tilde=w%*%y.miss
sigma.tilde=w%*%diag(sigma.miss)%*%t(w)
y.ash=ash(as.numeric(y.tilde),as.numeric(diag(sigma.tilde)))$result$PosteriorMean
y.hat=t(w)%*%y.ash
```



```{r,message=F}
plot(wd(y.miss,filter.number = 1,family='DaubExPhase'),main = 'Decomposition of Y with missing data set to 0')
plot(wd(y,filter.number = 1,family='DaubExPhase'),main='Decomposition with true Y, no missing data')
```

Let's look at $\tilde\sigma$:
```{r}
diag(sigma.tilde)
```
The 2nd-5th ones are from level 2, 6th-7th are from level 1 and the last one corresponds to level 0.

From the plot of coefficients, what we really want to set to 0 is 4th and 7th. But the 8th one also has huge variance. 

Can we manually choose what levels to shrink using large variance?

I don't think it's practical, especially when the length of sequence and number of missing data are large. Maybe the only criterion is whether it's visually appealing.

