---
title: "Poisson nugget effect simulation"
author: "Dongyue Xie"
date: "May 1, 2018"
output: html_document
---
Examine the performance of smash-gen(known $\sigma$) under different simulation settings.

# Algorithm

Let $X_t$ be a Poisson observation, $t=1,2,\dots,T$. 

1. Input $\sigma$ and initialize $m_t^{(0)}=\frac{\Sigma_{t=1}^T X_t}{T}$,  $Y_t^{(0)}=\log(m_t^{(0)})+\frac{X_t-m_t^{(0)}}{m_t^{(0)}}$ and $s_t^{2(0)}=\frac{1}{m_t^{(0)}}$ for $t=1,2,\dots,T$.
2. For $i=1,2,...$, iterate until convergence:
a. Fit $Y_t=\mu_t+N(0,\sigma^2)+N(0,s_t^2)$ using `smash.gaus` and obtain $\hat\mu_t$.
b. Update $m_t^{(i)}=\exp(\hat\mu_t)$, $Y_t^{(i)}=\log(m_t^{(i)})+\frac{X_t-m_t^{(i)}}{m_t^{(i)}}$, and $s_t^{2(i)}=\frac{1}{m_t^{(i)}}$

Convergence criteria: $||\mu_t^{(i)}-\mu_t^{(i-1)}||_2\leq \epsilon$.

```{r}
#' smash generaliation function

#' This function is for $Y_t=\mu_t+N(0,s_t^2)+N(0,\sigma^2)$ with known $s_t^2$ and $\sigma^2$.

#' @param x: a vector of observations
#' @param sigma: standard deviations, scalar.
#' @param family: choice of wavelet basis to be used, as in wavethresh.
#' @param niter: number of iterations for IRLS
#' @param tol: criterion to stop the iterations

smash.gen=function(x,sigma,family='DaubExPhase',niter=100,tol=1e-2){
  mu=c()
  s=c()
  mu=rbind(mu,rep(mean(x),length(x)))
  s=rbind(s,rep(1/mu[1],length(x)))
  y=log(mean(x))+(x-mean(x))/mean(x)
  for(i in 1:niter){
    mu.hat=smash.gaus(y,sigma=sigma+s[i,])
    mu=rbind(mu,mu.hat)
    #update m and s_t
    s=rbind(s,1/mu.hat)
    #update y
    mt=exp(mu.hat)
    y=log(mt)+(x-mt)/mt
    #y=log(mu.hat)+(x-mu.hat)/mu.hat
    if(norm(mu.hat-mu[i,],'2')<tol){
      break
    }
  }
  return(list(mu.hat=mu.hat,mu=mu,s=s))
}

```

Data generation by Poisson glm:

$\lambda_t=\exp(m_t+\epsilon_t)$, where $\epsilon_t\sim N(0,\sigma^2)$.

$X_t\sim Poi(\lambda_t)$.

```{r}
#' Simulation study comparing smash and smashgen

simu_study=function(m,sigma,seed=1234,
                    niter=100,family='DaubExPhase',tol=1e-2,
                    reflect=FALSE){
  set.seed(seed)
  lamda=exp(m+rnorm(length(m),0,sigma))
  x=rpois(length(m),lamda)
  #fit data
  smash.out=smash.poiss(x,reflect=FALSE)
  smash.gen.out=smash.gen(x,sigma=sigma,niter=niter,family = family,tol=tol)
  return(list(smash.out=smash.out,smash.gen.out=exp(smash.gen.out$mu.hat),smash.gen.est=smash.gen.out,x=x))
}

```

# Simulation 1: Constant trend Poisson nugget

$\sigma=0.01$

```{r}
library(smashr)
m=rep(3,256)
simu.out=simu_study(m,0.01)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", # places a legend at the appropriate place
       c("truth","smash-gen"), # puts text in the legend
       lty=c(1,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=0.1$

```{r}

simu.out=simu_study(m,0.1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```


$\sigma=0.5$

```{r}

simu.out=simu_study(m,0.5)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m),col='gray80')
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m),col='gray80')
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```
$\sigma=1$

```{r}

simu.out=simu_study(m,1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m),col='black')
legend("topleft", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m),col='black')
legend("topleft", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

# Simulation 2: Step trend 

$\sigma=0.01$

```{r}
m=c(rep(3,128), rep(5, 128), rep(6, 128), rep(4, 128))
simu.out=simu_study(m,0.01)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=0.1$

```{r}

simu.out=simu_study(m,0.1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=0.5$

```{r}

simu.out=simu_study(m,0.5)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=1$

```{r}

simu.out=simu_study(m,1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topleft", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topleft", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

# Simulation 3: Oscillating Poisson nugget

## Low Oscillating Poisson nugget

$\sigma=0.01$

```{r}
m=c()
for(k in 1:8){
    m=c(m, rep(1,16), rep(5, 16))
}
simu.out=simu_study(m,0.01)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))

```

$\sigma=0.1$

```{r}
simu.out=simu_study(m,0.1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))

```

$\sigma=0.5$

```{r}
simu.out=simu_study(m,0.5)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))

```

$\sigma=1$

```{r}
simu.out=simu_study(m,1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))

```

## Fast Oscillating Poisson nugget

$\sigma=0.01$

```{r}
m=c()
for(k in 1:32){
    m=c(m, c(1,5))
}
simu.out=simu_study(m,0.01)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
```

$\sigma=0.1$

```{r}
simu.out=simu_study(m,0.1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
```

$\sigma=0.5$

```{r}
simu.out=simu_study(m,0.5)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
```


$\sigma=1$

```{r}
simu.out=simu_study(m,1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '',main='smash-gen')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))


plot(simu.out$x,col = "gray80" ,ylab = '',main='smash')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
```

# Simulation 4: Polynomial curve Poisson nugget

$\sigma=0.01$

```{r}
m = seq(-1,1,length.out = 256)
m = m^3-2*m+1
simu.out=simu_study(m,0.01)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=0.1$

```{r}
m = seq(-1,1,length.out = 256)
m = m^3-2*m+1
simu.out=simu_study(m,0.1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=0.5$

```{r}
m = seq(-1,1,length.out = 256)
m = m^3-2*m+1
simu.out=simu_study(m,0.5)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

$\sigma=1$

```{r}
m = seq(-1,1,length.out = 256)
m = m^3-2*m+1
simu.out=simu_study(m,1)

#par(mfrow = c(1,2))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.gen.out, col = "red", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth","smash-gen"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black","red", "blue"))

plot(simu.out$x,col = "gray80" ,ylab = '')
lines(simu.out$smash.out, col = "blue", lwd = 2)
lines(exp(m))
legend("topright", 
       c("truth", "smash"), 
       lty=c(1,1), 
       lwd=c(1,1),
       cex = 1,
       col=c("black", "blue"))
```

# Summary

1. Generally, smash-gen gives more smooth fit under large nugget effect. But sometimes it seems that the iterative algorithm does not converge.
2. When the nugget effect is small, smash-gen may not perform well especially under oscillating Poisson nugget and polynomial curve.
