---
title: "Poisson to approximate binomial"
author: "Dongyue Xie"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

# Methods

Let $X\sim Binomial(n,p)$ then $E(X)=np, Var(X)=np(1-p)$. Poisson distribution is an approximation of binomial distribution when $n$ is large and $p$ is small. A rule of thumb is that $n\geq 20, p\leq 0.05$.

Derivation: Let $\lambda=np$

$$\frac{n!}{x!(n-x)!}p^x(1-p)^{n-x}=\frac{n(n-1)...(n-k+1)}{x!}(\lambda/n)^x(1-\lambda/n)^{n-x}\approx \frac{\lambda^x}{x!}(1-\lambda/n)^{n-x}$$ as $n\to \infty$. Since $lim_{n\to \infty}(1-\lambda/n)^{n}=e^{-\lambda}$ and $lim_{n\to \infty}(1-\lambda/n)^{-x}=1$, we have $$\frac{n!}{x!(n-x)!}p^x(1-p)^{n-x}\approx \frac{\lambda^x e^{-\lambda}}{x!}.$$

If we have binomial observation $X_t$ with $n_t$ and treat it as Poisson observation, we can do the following expansion: $$Y_t=\log(X_t)=\log(n_tp_t)+\frac{X_t-n_tp_t}{n_tp_t}=\log(n_t)+\log(p_t)+\frac{X_t-n_tp_t}{n_tp_t}.$$ This leads to $$Y_t-\log(n_t)=\log(p_t)+\frac{X_t-n_tp_t}{n_tp_t}.$$


# Experiments

We compare the performance of smashgen - binomial and smashgen - poi_binom, as well as Translation Invariant (TI) thresholding (Coifman and Donoho, 1995), which is one of the best methods in a large-scale simulation study in Antoniadis et al. (2001), and Ebayesthresh (Johnstone and Silverman, 2005b). 

For all experiments, T is set to be 256, nugget effect $\sigma=1$. The mean squared errors are reported and the plots are served as visual aids.

```{r}
library(smashrgen)
library(ggplot2)
library(EbayesThresh)

simu_study=function(p,sigma=1,ntri,nsimu=100,seed=12345,
                    niter=1,family='DaubExPhase',ashp=TRUE,verbose=FALSE,robust=FALSE,
                    tol=1e-2){
  set.seed(seed)
  smash.binom.err=c()
  smash.poibinom.err=c()
  ti.thresh.err=c()
  eb.thresh.err=c()
  n=length(p)
  target=exp(p)/(1+exp(p))
  for(k in 1:nsimu){
    ng=rnorm(n,0,sigma)
    m=exp(p+ng)
    q=m/(1+m)
    x=rbinom(n,ntri,q)
    #fit data
    smash.binom.out=smash_gen(x,dist_family = 'binomial',y_var_est='smashu',ntri=ntri)
    smash.poibinom.out=smash_gen(x,dist_family = 'poi_binom',y_var_est='smashu',ntri=ntri)
    ti.thresh.out=ti.thresh(x/ntri,method='rmad')
    eb.thresh.out=waveti.ebayes(x/ntri)
    #errors
    smash.binom.err[k]=mse(target,smash.binom.out)
    smash.poibinom.err[k]=mse(target,smash.poibinom.out)
    ti.thresh.err[k]=mse(target,ti.thresh.out)
    eb.thresh.err[k]=mse(target,eb.thresh.out)
  }
  return(list(est=list(smash.binom.out=smash.binom.out,smash.poibinom.out=smash.poibinom.out, ti.thresh.out=ti.thresh.out,eb.thresh.out=eb.thresh.out,x=x),
              err=data.frame(smash.binom=smash.binom.err,smash.poibinom=smash.poibinom.err, ti.thresh=ti.thresh.err,eb.thresh=eb.thresh.err)))
}

waveti.ebayes = function(x, filter.number = 10, family = "DaubLeAsymm", min.level = 3) {
    n = length(x)
    J = log2(n)
    x.w = wd(x, filter.number, family, type = "station")
    for (j in min.level:(J - 1)) {
        x.pm = ebayesthresh(accessD(x.w, j))
        x.w = putD(x.w, j, x.pm)
    }
    mu.est = AvBasis(convert(x.w))
    return(mu.est)
}
```

## Constant trend

### $ntri$ large, $p$ large

The number of trials are generated from a Poisson distribution with $\lambda=50$. $p$ is around 0.8.

```{r}
n=256
p=rep(1.5,n)
set.seed(111)
ntri=rpois(n,50)
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)


```

### $ntri$ small, $p$ large

The number of trials are generated from a Poisson distribution with $\lambda=8$. $p$ is around 0.8.

```{r}
n=256
p=rep(1.5,n)
set.seed(111)
ntri=rpois(n,5)+1
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)


```

### $ntri$ large, $p$ small

The number of trials are generated from a Poisson distribution with $\lambda=50$. $p$ is around 0.05.

```{r}
n=256
p=rep(-3,n)
set.seed(111)
ntri=rpois(n,50)
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
ggplot(df2gg(result$err[,1:2]),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)


```

As expected, when $n$ is large and $p$ is small, Poisson distribution is a good approximation to binomial distribution. 

### $ntri$ small, $p$ small

The number of trials are generated from a Poisson distribution with $\lambda=5$. $p$ is around 0.05.

```{r}
n=256
p=rep(-3,n)
set.seed(111)
ntri=rpois(n,5)+1
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)


```

## Steps

# Large $n$

```{r}
p=c(rep(-2,64), rep(0, 64), rep(2, 64), rep(-2, 64))
set.seed(111)
ntri=rpois(256,50)
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)
```

# Small $n$

```{r}
p=c(rep(-2,64), rep(0, 64), rep(2, 64), rep(-2, 64))
set.seed(111)
ntri=rpois(256,5)+1
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)
```

## Bumps

### Large $n$.

```{r}
m=seq(0,1,length.out = 256)
h = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
w = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005,0.008,0.005)
t=c(.1,.13,.15,.23,.25,.4,.44,.65,.76,.78,.81)
f = c()
for(i in 1:length(m)){
  f[i]=sum(h*(1+((m[i]-t)/w)^4)^(-1))
}
p=f-3

set.seed(111)
ntri=rpois(256,50)
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)

```

### Small $n$

```{r}
set.seed(111)
ntri=rpois(256,5)+1
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)
```

## Spike mean

### Large $n$

```{r}
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 256
t = 1:n/n
p = spike.f(t)*2-2

set.seed(111)
ntri=rpois(256,50)
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)
```


### Small $n$

```{r}
set.seed(111)
ntri=rpois(256,5)+1
result=simu_study(p,ntri=ntri)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')

par(mfrow=c(2,2))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.binom.out,col=4,lwd=2)
legend("bottomright", # places a legend at the appropriate place
       c("truth","smash-binom"), # puts text in the legend
       lty=c(2,1), # gives the legend appropriate symbols (lines)
       lwd=c(1,2),
       cex = 1,
       col=c("black","blue"))
plot(result$est$x/ntri,col='gray80',ylab='',main='smash-poi_binom')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$smash.poibinom.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='TI thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$ti.thresh.out,col=4,lwd=2)
plot(result$est$x/ntri,col='gray80',ylab='',main='EBayes thresh')
lines(exp(p)/(1+exp(p)),lty=2)
lines(result$est$eb.thresh.out,col=4,lwd=2)
```
