---
title: "Poisson spike issues"
author: "Dongyue Xie"
date: "2018-10-07"
output: 
  workflowr::wflow_html:
  code_folding: hide
---

# The problem

Model: $X\sim Poi(\mu)$ and define $y=\log m+\frac{x-m}{m}$

Previously, we used $m=$ash posterior mean of x. The problem is that shrinkage effect is too strong that for large observations $x$, the approximated normal data points are too large. Hence, after taking exponential of estimated normal mean, the estimation 'blow up'. 

Now we try to do Taylor series expansion around $\log x$(MLE) for non-zero $x$ and around ash posterior mean around zero $x$. In other words, now $m=$ash posterior mean of zero $x$s and $m=x$ for non-zero $x$s. 

# Spike mean function

```{r,message=FALSE,warning=FALSE,fig.width=12,fig.height=10}
library(ashr)
library(smashrgen)
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 256
t = 1:n/n
m = spike.f(t)

m=m*2+0.1
range(m)

sig=0

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)

x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='n=256,nugget=0')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))
#################

sig=0.1

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)


x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

plot(x,col='grey80',ylab='',xlab='',main='n=256,nugget=0.1')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))

#################

sig=1

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)

x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

plot(x,col='grey80',ylab='',xlab='',main='n=256,nugget=1')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))

plot(x,col='grey80',ylab='',xlab='',main='Previous verison using ash posterior mean, nugget=1')
lines(m,col='grey60')
lines(smash_gen_lite(x))
legend('topleft',c('data','true mean','fit'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey60',1))
```


```{r,fig.width=12,fig.height=10}
n = 512
t = 1:n/n
m = spike.f(t)

m=m*2+0.1
range(m)

sig=0

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='n=512,nugget=0')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))
#################

sig=0.1

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

plot(x,col='grey80',ylab='',xlab='',main='n=512,nugget=0.1')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))

#################

sig=1

set.seed(12345)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



x.ash=ash(rep(0,n),1,lik=lik_pois(x))$result$PosteriorMean
m.hat=x.ash
m.hat[which(x!=0)]=(x[which(x!=0)])
y=log(m.hat)+(x-m.hat)/m.hat
m.tilde=exp(smash.gaus(y,sigma = sqrt(sig^2+1/m.hat)))
m.tilde2=exp(smash.gaus(y))

plot(x,col='grey80',ylab='',xlab='',main='n=512,nugget=1')
lines(m,col='grey60')
lines(m.tilde,col=2)
lines(m.tilde2,col=4)

legend('topleft',c('data','true mean','smashgen-known var','smashgen-unknown var'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey60',2,4))

plot(x,col='grey80',ylab='',xlab='',main='Previous verison using ash posterior mean, nugget=1')
lines(m,col='grey60')
lines(smash_gen_lite(x))
legend('topleft',c('data','true mean','fit'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey60',1))
```
