---
title: "Poisson nugget effect, unknown variance"
author: "Dongyue Xie"
date: "May 13, 2018"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

Simulations of Poisson nugget effect(unkown).

Previously, we have studies the [methods](sigma.html) to estimate unknown $\sigma$ in the model $Y_t=\mu_t+N(0,\sigma^2)+N(0,s_t^2)$. Here, we apply mle and smashu methods and compare them with smash as well as smashgen with known $sigma$. The measure of accuracy is mean square error. Plots are also given as visual aid.

```{r,warning=FALSE}
library(smashrgen)
library(ggplot2)
```

```{r,warning=FALSE}
#' Simulation study comparing smash and smashgen

simu_study=function(m,sigma,nsimu=100,seed=12345,
                    niter=1,family='DaubExPhase',ashp=TRUE,verbose=FALSE,robust=FALSE,
                    tol=1e-2){
  set.seed(seed)
  smash.err=c()
  smashgen.err=c()
  smashgen.smashu.err=c()
  smashgen.mle.err=c()
  for(k in 1:nsimu){
    lamda=exp(m+rnorm(length(m),0,sigma))
    x=rpois(length(m),lamda)
    #fit data
    smash.out=smash.poiss(x)
    smashgen.out=smash_gen(x,dist_family = 'poisson',sigma = sigma)
    smashu.out=smash_gen(x,dist_family = 'poisson',y_var_est = 'smashu')
    mle.out=smash_gen(x,dist_family = 'poisson',y_var_est = 'mle')
    smash.err[k]=mse(exp(m),smash.out)
    smashgen.err[k]=mse(exp(m),smashgen.out)
    smashgen.smashu.err[k]=mse(exp(m),smashu.out)
    smashgen.mle.err[k]=mse(exp(m),mle.out)
  }
  return(list(est=list(smash.out=smash.out,smashgen.out=smashgen.out,smashu.out=smashu.out,mle.out=mle.out,x=x),err=data.frame(smash=smash.err,smashgen=smashgen.err,
              smashgen.smashu=smashgen.smashu.err,smashgen.mle=smashgen.mle.err)))
}
```

# Simulation 1: Constant trend Poisson nugget

## $\sigma=0.1$

```{r,warning=FALSE}
m=rep(1,128)
result=simu_study(m,0.1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$

```{r,warning=FALSE}
m=rep(1,128)
result=simu_study(m,1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=2$

```{r,warning=FALSE}
m=rep(1,128)
result=simu_study(m,2)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

# Simulation 2: Step trend 

## $\sigma=0.1$

```{r,warning=FALSE}
m=c(rep(3,128), rep(5, 128), rep(6, 128), rep(3, 128))
result=simu_study(m,0.1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$

```{r,warning=FALSE}
m=c(rep(3,128), rep(5, 128), rep(6, 128), rep(3, 128))
result=simu_study(m,1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

# Simulation 3: Bumps

## $\sigma=0.1$
```{r,warning=FALSE}
m=seq(0,1,length.out = 256)
h = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
w = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005,0.008,0.005)
t=c(.1,.13,.15,.23,.25,.4,.44,.65,.76,.78,.81)
f = c()
for(i in 1:length(m)){
  f[i]=sum(h*(1+((m[i]-t)/w)^4)^(-1))
}
m=f
result=simu_study(m,0.1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$
```{r,warning=FALSE}
result=simu_study(m,1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

# Simulation 4: Spike mean

## $\sigma=0.1$
```{r,warning=FALSE}
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 256
t = 1:n/n
m = spike.f(t)

result=simu_study(m,0.1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```

## $\sigma=1$
```{r,warning=FALSE}
result=simu_study(m,1)
par(mfrow=c(2,2))
plot(result$est$x,col='gray80',ylab='',main='smash')
lines(exp(m),col=1)
lines(result$est$smash.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: known sigma')
lines(exp(m),col=1)
lines(result$est$smashgen.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: (sigma^2+s_t^2) from smash')
lines(exp(m),col=1)
lines(result$est$smashu.out,col=4)
plot(result$est$x,col='gray80',ylab='',main='smashgen: sigma from mle')
lines(exp(m),col=1)
lines(result$est$mle.out,col=4)

mean(result$err$smash)
mean(result$err$smashgen)
mean(result$err$smashgen.smashu)
mean(result$err$smashgen.mle)

ggplot(df2gg(result$err),aes(x=method,y=MSE))+geom_boxplot(aes(fill=method))+labs(x='')
```
