---
title: "Poisson seqence and wave functions"
author: "Dongyue Xie"
date: "May 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F,message = F)
```

We have shown that smashgen-Poisson outperforms smash when smoothing Poisson data with nugget effect. One natural question is: is this true for Poisson data(no nugget effect)? 

One limitation of `smash.pois` is that it can only use an analogue
of the Haar wavelet transform to the Poisson data. So it may lose power when dealing with wavelet whose signal is better captured by more complex basis functions.

In this analysis, we try to address the above two questions.

We have seen in the previous [analysis](wavebasis.html) that Symmlet 8 gives better decomposition for smoother functions, while for functions with spike or sharp changes, the two basis have similar results. 

```{r}
simu_study_basis=function(mu,nsimu=100,seed=1234,niter=1,robust=F){
  n=length(mu)
  set.seed(seed)
  smash.err=c()
  gen.haar.err=c()
  gen.sym.err=c()
  for(iter in 1:nsimu){
    y=rpois(n,mu)
    smash.out=smash.poiss(y)
    gen.haar.out=smash_gen(y,niter = niter,robust=robust)
    gen.sym.out=smash_gen(y,wave_family = 'DaubLeAsymm',filter.number = 8,niter=niter,robust=robust)
    smash.err[iter]=mse(smash.out,mu)
    gen.haar.err[iter]=mse(gen.haar.out,mu)
    gen.sym.err[iter]=mse(gen.sym.out,mu)
  }
  return(list(est=data.frame(smash=smash.out,smashgen.haar=gen.haar.out,smashgen.sym=gen.sym.out),err=data.frame(smash=smash.err,smashgen.haar=gen.haar.err,smashgen.sym=gen.sym.err)))
}

```

# HeaviSine

range of $\mu$ roughly $(0.3,7)$.
```{r}
library(smashrgen)

mu=DJ.EX(256,signal = 2)$heavi
mu=mu-min(mu)+0.3
plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(0.1,67)$.
```{r}
mu=DJ.EX(256,signal = 20)$heavi
mu=mu-min(mu)+0.1
#plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

# Doppler
range of $\mu$ roughly $(0.1,7)$.
```{r}
mu=DJ.EX(256,signal = 2)$doppler
mu=mu-min(mu)+0.1
plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(0.1,70)$.
```{r}
mu=DJ.EX(256,signal = 20)$doppler
mu=mu-min(mu)+0.1
#plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

# Parabola

range of $\mu$ roughly $(0.1,6)$.
```{r}
r=function(x,c){return((x-c)^2*(x>c)*(x<=1))}
f=function(x){return(0.8 − 30*r(x,0.1) + 60*r(x, 0.2) − 30*r(x, 0.3) +
500*r(x, 0.35) − 1000*r(x, 0.37) + 1000*r(x, 0.41) − 500*r(x, 0.43) +
7.5*r(x, 0.5) − 15*r(x, 0.7) + 7.5*r(x, 0.9))}
mu=f(1:256/256)
mu=mu*10-1.9
plot(mu,type = 'l')
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(1,73)$.
```{r}
r=function(x,c){return((x-c)^2*(x>c)*(x<=1))}
f=function(x){return(0.8 − 30*r(x,0.1) + 60*r(x, 0.2) − 30*r(x, 0.3) +
500*r(x, 0.35) − 1000*r(x, 0.37) + 1000*r(x, 0.41) − 500*r(x, 0.43) +
7.5*r(x, 0.5) − 15*r(x, 0.7) + 7.5*r(x, 0.9))}
mu=f(1:256/256)
mu=mu*120-23
#plot(mu,type = 'l')
result=simu_study_basis(mu)
boxplot(result$err)
```

# Step

range of $\mu$ roughly $(1,6)$.
```{r}
mu=c(rep(1,64), rep(3, 64), rep(6, 64), rep(1, 64))
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(1,80)$.
```{r}
mu=c(rep(1,64), rep(30, 64), rep(80, 64), rep(1, 64))
result=simu_study_basis(mu)
boxplot(result$err)
```

# Wave

range of $\mu$ roughly $(0.25,6)$.
```{r}
f=function(x){return(0.5 + 0.2*cos(4*pi*x) + 0.1*cos(24*pi*x))}
mu=f(1:256/256)
mu=mu*10-2
plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(0.3,75)$.
```{r}
mu=f(1:256/256)
mu=mu*130-29
#plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

# Time shifted sine

range of $\mu$ roughly $(0.1,6)$.

```{r}
g=function(x){return((1 − cos(pi*x))/2)}
f=function(x){return(0.3*sin(3*pi*(g(g(g(g(x)))) + x) + 0.5))}
mu=f(1:256/256)
mu=mu*10+3.1
plot(mu,type='l')
result=simu_study_basis(mu)
boxplot(result$err)
```

range of $\mu$ roughly $(0.1,120)$.
```{r}
mu=f(1:256/256)
mu=mu*200+60.1
#plot(mu)
result=simu_study_basis(mu)
boxplot(result$err)
```

