---
title: "Shrink R squared - examples"
author: "Dongyue Xie"
date: "2019-01-21"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

For the method used in these examples, see [here](r2.html)

1. True $R^2$ is defined as $R^2=\frac{var(X\beta)}{var(y)}=\frac{var(y)-\sigma^2}{var(y)}=1-\frac{\sigma^2}{var(y)}=1-\frac{\sigma^2}{\sigma^2+var(X\beta)}$

2. Ajusted R^2: $1-\frac{\sum(y_i-\hat y_i)^2/(n-p-1)}{\sum(y_i-\bar y)^2/(n-1)}$

3. Shrunk adjusted R^2: use `fash` shrinking $fash.output=\log(\frac{\sum(y_i-\hat y_i)^2/(n-p-1)}{\sum(y_i-\bar y)^2/(n-1)})$ then shrunk adjusted R^2 is $1-\exp(fash.output)$

4. Shrunk R^2: use `fash` shrinking $fash.output=\log(\frac{\sum(y_i-\hat y_i)^2/(n-1)}{\sum(y_i-\bar y)^2/(n-1)})$ then shrunk adjusted R^2 is $1-\exp(fash.output)$

5. Another Shrunk R^2: shrink all $\betas$ using `ash`, obtain posterior means then calculate $\hat\sigma^2$ then obtain $\frac{var(X\hat\beta)}{var(\hat\beta)+\hat\sigma^2}$.

(note: this is the old idea which introduces bias to R^2 when multiplying $\frac{n-p-1}{n-1}$ so I discard this method.)(Shrunk R^2 = $1 - \exp(fash.output)*\frac{n-p-1}{n-1}$, because $R^2=1-\frac{\sum(y_i-\hat y_i)^2}{\sum(y_i-\bar y)^2}=1-\frac{\sum(y_i-\hat y_i)^2/(n-p-1)}{\sum(y_i-\bar y)^2/(n-1)}*\frac{n-p-1}{n-1}=1-(1-adjR^2)*\frac{n-p-1}{n-1}$)

## R Function

R function for shrinking adjusted R/ R squared:

```{r}
library(ashr)
#'@param R2: R squared from linear regression model fit
#'@param n: sample size
#'@param p: the number of covariates
#'@output shrunk R squared.

ash_ar2=function(R2,n,p){
  df1=n-p-1
  df2=n-1
  log.ratio=log((1-R2)/(df1)*(df2))
  shrink.log.ratio=ash(log.ratio,1,lik=lik_logF(df1=df1,df2=df2))$result$PosteriorMean
  ar2=1-exp(shrink.log.ratio)
  return(ar2)
}


ash_r2=function(R2,n,p){
  df1=n-1
  df2=n-1
  log.ratio=log(1-R2)
  shrink.log.ratio=ash(log.ratio,1,lik=lik_logF(df1=df1,df2=df2))$result$PosteriorMean
  r2=1-exp(shrink.log.ratio)
  return(r2)
}


```

## Compare Shrunk $R^2$ with True $R2$.

Assume linear model $y=X\beta+\epsilon$ where $\epsilon\sim N(0,\sigma^2I)$

1. n=100, p=5. Each cordinate of $\beta$ ranges from 0 to 1, for example $\beta=(0,0,0,0,0)$,...,$\beta=(0.1,0.1,0.1,0.1,0.1)$,..., $\beta=(1,1,1,1,1)$ etc.  


**If I generate X from Uniform(0,1), then fash shirnks all $R^2$ to 0. If generate X from Uniform(0,2), then it does not**

Uniform(0,1):

```{r}
set.seed(1234)

n=100
p=5
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=var(X%*%beta)/(1+var(X%*%beta))
}
R2s=ash_r2(R2,n,p)
R2as=ash_ar2(R2,n,p)
plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
lines(beta.list,R2as,col=3)
lines(beta.list,R2s,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2as,col=3)
lines(trueR2,R2s,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash'),lty=c(1,1,1,1),col=c(1,2,3,4))

```


Uniform(0,3):

```{r}
set.seed(1234)

n=100
p=5
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,3),n,p)
for (i in 1:length(beta.list)) {
  
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=var(X%*%beta)/(1+var(X%*%beta))
}
R2s=ash_r2(R2,n,p)
R2as=ash_ar2(R2,n,p)
plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
lines(beta.list,R2as,col=3)
lines(beta.list,R2s,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2as,col=3)
lines(trueR2,R2s,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash'),lty=c(1,1,1,1),col=c(1,2,3,4))

```

Uniform(0,5):

```{r}
set.seed(1234)

n=100
p=5
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=var(X%*%beta)/(1+var(X%*%beta))
}
R2s=ash_r2(R2,n,p)
R2as=ash_ar2(R2,n,p)
plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
lines(beta.list,R2as,col=3)
lines(beta.list,R2s,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2as,col=3)
lines(trueR2,R2s,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash'),lty=c(1,1,1,1),col=c(1,2,3,4))

```

2. Increase $p$ to 20.

**This time, if I generate X from Uniform(0,1), then fash does not shirnk all $R^2$ to 0.**

Uniform(0,1):

```{r}
set.seed(1234)

n=100
p=20
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=var(X%*%beta)/(1+var(X%*%beta))
}
R2s=ash_r2(R2,n,p)
R2as=ash_ar2(R2,n,p)
plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
lines(beta.list,R2as,col=3)
lines(beta.list,R2s,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2as,col=3)
lines(trueR2,R2s,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash'),lty=c(1,1,1,1),col=c(1,2,3,4))

```

2. $n=30, p=3$. 

**This time, I have to generate X from at least Uniform(0,3) to avoid over-shrinkage of fash. Here, I tried Uniform(0,5)**

Uniform(0,5):

```{r}
set.seed(1234)

n=30
p=3
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=var(X%*%beta)/(1+var(X%*%beta))
}
R2s=ash_r2(R2,n,p)
R2as=ash_ar2(R2,n,p)
plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
lines(beta.list,R2as,col=3)
lines(beta.list,R2s,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2as,col=3)
lines(trueR2,R2s,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','Adjusted R^2 fash','R^2 fash'),lty=c(1,1,1,1),col=c(1,2,3,4))

```


## Compare fash and corshrink

### 1-d case

1. $n=100,p=1$

Uniform(0,5):

```{r}
library(CorShrink)
set.seed(1234)

n=100
p=1
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X*beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))




```

2. $n=30,p=1$

Uniform(0,5):

```{r}
set.seed(1234)

n=30
p=1
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X*beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))

```


### Multiple regression

1. $n=100, p=5$

Uniform(0,2):

```{r}
set.seed(1234)

n=100
p=5
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,2),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))

```

2. $n=100,p=20$

Uniform(0,1):

```{r}
set.seed(1234)

n=100
p=20
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))

```


$n=100,p=50$

Uniform(0,1):

```{r}
set.seed(1234)

n=100
p=50
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))

```

3. $n=30, p=3$

Uniform(0,5):

```{r}
set.seed(1234)

n=30
p=3
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,5,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}
R2.fash=ash_r2(R2,n,p)
R2a.fash=ash_ar2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2


plot(beta.list,R2,ylim=range(c(R2,R2a,R2s,R2as,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2a,col=2)
#lines(beta.list,R2a.fash,col=3)
lines(beta.list,R2.fash,col=3)
lines(beta.list,R2.cor,col=4)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink','True R^2'),lty=c(1,1,1,1,1),col=c(1,2,3,4,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2a,R2s,R2as,trueR2)))
lines(trueR2,R2a,col=2)
lines(trueR2,R2.fash,col=3)
lines(trueR2,R2.cor,col=4)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','Adjusted R^2','R^2 fash','R^2 CorShrink'),lty=c(1,1,1,1),col=c(1,2,3,4))

```


# Summary1

1. When generating X from Uniform(0,1), $var(X\beta)$ is small and `fash` can shrink all $R^2$ to 0. This happens when $n,p$ are small. If $p=20$, then this does not happen.

2. CorShrink does not shrink $R^2$. I can not really tell the difference from plots between Corshrink $R^2$ and $R&2$.

3. Adjusted $R^2$ is a good shrinkage estimator of $R^2$.

## Sign of correlations

Randomize signs of $R$ and see if corshrink gives the same results.

Random sample n/2 $R^2$s and set the sign of $R$ to negative.

$n=100,p=1$

Uniform(0,1):

```{r}
library(CorShrink)
set.seed(1234)

n=100
p=1
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X*beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}

R2.cor=(CorShrinkVector(sqrt(R2),n))^2
idx=sample(1:100,50)
R=sqrt(R2)
R[idx]=-R[idx]
R2.cor.sign=(CorShrinkVector(R,n))^2

plot(beta.list,R2,ylim=range(c(R2,R2.cor.sign,R2.cor,trueR2)),main='',xlab='beta',ylab='',type='l')
lines(beta.list,R2.cor,col=2)
lines(beta.list,R2.cor.sign,col=3)
lines(beta.list,trueR2,col='grey80')
abline(h=0,lty=2)
legend('topleft',c('R^2','R^2 CorShrink','R^2 CorShrink Random Sign','True R^2'),lty=c(1,1,1,1),col=c(1,2,3,'grey80'))

plot(trueR2,R2,type='l',ylim=range(c(R2,R2.cor.sign,R2.cor,trueR2)))
lines(trueR2,R2.cor,col=2)
lines(trueR2,R2.cor.sign,col=3)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('R^2','R^2 CorShrink','R^2 CorShrink Random Sign'),lty=c(1,1,1),col=c(1,2,3))




```


So signs do not matter.


## Estimates of g

### Example 0

X from Uniform(0,5) and $n=100,p=1$

```{r}
set.seed(1234)
n=100
p=1
R2=c()
R2a=c()
trueR2=c()
beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}

R2.fash=ash_r2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2

plot(trueR2,R2.fash,type='l',ylim=range(c(R2.fash,R2.cor,trueR2)),ylab = '')
lines(trueR2,R2.cor,col=2)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('fash','Corshrink','True R2'),lty=c(1,1,1),col=c(1,2,'grey80'))

```

fash:

```{r}
log.ratio=log(1-R2)
ash.fit=ash(log.ratio,1,lik=lik_logF(df1=n-1,df2=n-1))
ash.fit$fitted_g
```

Fitted g concentrates at $0.4*\delta_0+0.6*Uniform(-1.07,1.07)$

Corshrink:

```{r}
R=sqrt(R2)
corvec=R
corvec_trans = 0.5 * log((1 + corvec)/(1 - corvec))
corvec_trans_sd = rep(sqrt(1/(n - 1) + 2/(n - 
            1)^2), length(corvec_trans))

ash.control=list()
ash.control.default = list(pointmass = TRUE, mixcompdist = "normal", 
        nullweight = 10, fixg = FALSE, mode = 0, optmethod = "mixEM", 
        prior = "nullbiased", gridmult = sqrt(2), outputlevel = 2, 
        alpha = 0, df = NULL, control = list(K = 1, method = 3, 
            square = TRUE, step.min0 = 1, step.max0 = 1, mstep = 4, 
            kr = 1, objfn.inc = 1, tol = 1e-05, 
            trace = FALSE))
ash.control <- utils::modifyList(ash.control.default, ash.control)
    

fit = do.call(ashr::ash, append(list(betahat = corvec_trans, 
        sebetahat = corvec_trans_sd), ash.control))
fit$fitted_g
```

Fitted g concentrates at $0.12*\delta_0+0.5*N(0,0.62^2)+0.37*N(0,0.87^2)$


### Example 1

X from Uniform(0,1) and $n=100,p=5$

```{r}
set.seed(1234)
n=100
p=5
R2=c()
R2a=c()
trueR2=c()
beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,1),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}

R2.fash=ash_r2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2

plot(trueR2,R2.fash,type='l',ylim=range(c(R2.fash,R2.cor,trueR2)),ylab = '')
lines(trueR2,R2.cor,col=2)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('fash','Corshrink','True R2'),lty=c(1,1,1),col=c(1,2,'grey80'))

```

fash:

```{r}
log.ratio=log(1-R2)
ash.fit=ash(log.ratio,1,lik=lik_logF(df1=n-1,df2=n-1))
ash.fit$fitted_g
```

Fitted g is at point mass at 0.

Corshrink:

```{r}
R=sqrt(R2)
corvec=R
corvec_trans = 0.5 * log((1 + corvec)/(1 - corvec))
corvec_trans_sd = rep(sqrt(1/(n - 1) + 2/(n - 
            1)^2), length(corvec_trans))

ash.control=list()
ash.control.default = list(pointmass = TRUE, mixcompdist = "normal", 
        nullweight = 10, fixg = FALSE, mode = 0, optmethod = "mixEM", 
        prior = "nullbiased", gridmult = sqrt(2), outputlevel = 2, 
        alpha = 0, df = NULL, control = list(K = 1, method = 3, 
            square = TRUE, step.min0 = 1, step.max0 = 1, mstep = 4, 
            kr = 1, objfn.inc = 1, tol = 1e-05, 
            trace = FALSE))
ash.control <- utils::modifyList(ash.control.default, ash.control)
    

fit = do.call(ashr::ash, append(list(betahat = corvec_trans, 
        sebetahat = corvec_trans_sd), ash.control))
fit$fitted_g
```

Fitted g concentrates at $0.9*N(0,0.41^2)$

### Example 2

X from Uniform(0,5) and $n=100,p=5$

```{r}
set.seed(1234)
n=100
p=5
R2=c()
R2a=c()
trueR2=c()

beta.list=seq(0,1,length.out = 100)
X=matrix(runif(n*(p),0,5),n,p)
for (i in 1:length(beta.list)) {
  beta=rep(beta.list[i],p)
  y=X%*%beta+rnorm(n)
  datax=data.frame(X=X,y=y)
  mod=lm(y~.,datax)
  mod.sy=summary(mod)
  R2[i]=mod.sy$r.squared
  R2a[i]=mod.sy$adj.r.squared
  trueR2[i]=1-(1)/(1+var(X%*%beta))
}

R2.fash=ash_r2(R2,n,p)
R2.cor=(CorShrinkVector(sqrt(R2),n))^2

plot(trueR2,R2.fash,type='l',ylim=range(c(R2.fash,R2.cor,trueR2)),ylab='')
lines(trueR2,R2.cor,col=2)
lines(trueR2,trueR2,col='grey80')
legend('topleft',c('fash','Corshrink','True R2'),lty=c(1,1,1),col=c(1,2,'grey80'))

```

fash:

```{r}
log.ratio=log(1-R2)
ash.fit=ash(log.ratio,1,lik=lik_logF(df1=n-1,df2=n-1))
ash.fit$fitted_g
```

Fitted g concentrates at $0.19*\delta_0+0.81*Uniform(-2.38,2.38)$.

Corshrink:

```{r}
R=sqrt(R2)
corvec=R
corvec_trans = 0.5 * log((1 + corvec)/(1 - corvec))
corvec_trans_sd = rep(sqrt(1/(n - 1) + 2/(n - 
            1)^2), length(corvec_trans))

ash.control=list()
ash.control.default = list(pointmass = TRUE, mixcompdist = "normal", 
        nullweight = 10, fixg = FALSE, mode = 0, optmethod = "mixEM", 
        prior = "nullbiased", gridmult = sqrt(2), outputlevel = 2, 
        alpha = 0, df = NULL, control = list(K = 1, method = 3, 
            square = TRUE, step.min0 = 1, step.max0 = 1, mstep = 4, 
            kr = 1, objfn.inc = 1, tol = 1e-05, 
            trace = FALSE))
ash.control <- utils::modifyList(ash.control.default, ash.control)
    

fit = do.call(ashr::ash, append(list(betahat = corvec_trans, 
        sebetahat = corvec_trans_sd), ash.control))
fit$fitted_g
```

Fitted g concentrates at $0.91*N(0,1.39^2)$


# Summary2

Fash uses a mixture of point mass and uniform distributions as prior while CorShrink uses a mixture of point mass and normal distirbutions. Fash in these examples put more weights on point mass in fitted g than CorShrink.


# Thoughts

1. Applying `fash` to shrink $R^2$ replies on F distribution, which is from the ratio of two variances. F distribution replies on normal assumption and independence of two normal populations. However, $var(\sigma^2)$ and $var(y)=var(X\beta)+var(\sigma^2)$ are not independent. So rigourously speaking, using `fash` is not appropriate here. 

2. `Corshirnk` depends on Fisher transformation which has bivariate normal assumption. Since $R^2=r^2_{y,\hat y}$, we can apply `Corshirnk` if $y,\hat y$ is bivariate normal distributed. By saying $y,\hat y$ is bivariate normal distributed, I mean $y,\hat y$ are $n$ i.i.d samples from a bivariate normal distribution. However, this can hardly be true because $\hat y=Hy$ where $H=X(X^TX)^{-1}X^T$, so the $n$ samples $y,\hat y$ are not independently generated.

3. From the examples above, adjusted $R^2$ is a good estimate of true $R^2$. It gives estimate close to ture $R^2$ which can be seen from the True $R^2$ - estimated $R^2$ plot. It's pitfall it that it's no longer necessatily positive - it can be negative.
