---
title: "Functional Data analysis with smashgen"
author: "Dongyue Xie"
date: "2019-02-27"
output: 
  workflowr::wflow_html:
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

# Introduction

Assume we obseve $Y\in R^{N*T}$: N curves(count) and each has length T from N individuals. We also obseve a $X\in R^{N*p}$ data matrix whose columns are covariates. We can model $g(Y)=XB+E$, where $B\in R^{p*T}$ is coefficient matrix, whose rows are smooth curves, and $E\in R^{N*T}$ is a random error matrix.

Model framework:

1. Apply log or vst transformation to $Y$ and obtain $\tilde Y=XB+E_1+E_2$, where $E_1$ is nugget effect matrix and $E_2$ is variance matrix resulted from transformation.
2. Apply DWT: $\tilde Y W=XBW+E_1W+E_2W$. Write it as $Y^*=XB^*+E_1*+E_2^*$
3. Fit weighted linear regression for each $t$, $t=1,2,3,...,T$ and obtain $\hat B^*$. 
4. Apply `ash` to each row of $\hat B^*$ and obtain new $\hat B^*$
5. Apply inverse DWT and obtain $\hat B$.


```{r}
wavelet_fda=function(Y,X,sigma=1,ca=0,trans='log',filter.number=1,family='DaubExPhase'){
  
  N=nrow(Y)
  Tt=ncol(Y)
  p=ncol(X)
  if(length(sigma)==1){sigma=rep(sigma,N)}
  W=GenW(n=Tt,filter.number=filter.number,family=family)
  if(trans=='log'){
    Y.t=matrix(nrow=N,ncol=Tt)
    set2=matrix(nrow=N,ncol=Tt)
    for (n in 1:N) {
      x=Y[n,]
      x.ash=ash(rep(0,Tt),1,lik=lik_pois(x))$result$PosteriorMean
      m.hat=x.ash
      m.hat[which(x!=0)]=(x[which(x!=0)])
      ys=log(m.hat)+(x-m.hat)/m.hat
      st2=1/(m.hat)
      
      Y.t[n,]=W%*%ys
      set2[n,]=diag(W%*%diag(sigma[n]^2+st2)%*%t(W))
    }

  }
  if(trans=='vst'){
    Y.t=t(apply(Y, 1, function(x){W%*%sqrt(x+ca)}))
    set2=matrix(nrow=N,ncol=Tt)
    for (n in 1:N) {
      st2=1/4
      set2[n,]=diag(W%*%diag(rep(sigma[n]^2+st2,Tt))%*%t(W))
    }
  }
  
  beta.hat=matrix(nrow=p+1,ncol=Tt)
  beta.hat.se=matrix(nrow=p+1,ncol=Tt)
  
  for (t in 1:Tt) {
    wls.fit=lm(y~.,data.frame(y=Y.t[,t],x=X),weights = 1/set2[,t])
    beta.hat[,t]=coefficients(wls.fit)
    smy=summary(wls.fit)
    beta.hat.se[,t]=smy$coefficients[,2]
  }
  
  beta.hat.shrink=matrix(nrow=p+1,ncol=Tt)
  for (j in 1:(p+1)) {
    beta.hat.shrink[j,]=ashr::ash(beta.hat[j,],beta.hat.se[j,])$result$PosteriorMean
  }
  return(beta.hat.shrink%*%W)
}



```

# Simulation

Let N=100, p=3(includes $1_N$).

```{r}
library(wavethresh)
library(ashr)
Tt=512
  beta1=c(rep(2,Tt/4), rep(5, Tt/4), rep(6, Tt/4), rep(2, Tt/4))
beta1=5*beta1/sqrt(norm(beta1,'2'))

r=function(x,c){return((x-c)^2*(x>c)*(x<=1))}
f=function(x){return(0.8 - 30*r(x,0.1) + 60*r(x, 0.2) - 30*r(x, 0.3) +
                       500*r(x, 0.35) -1000*r(x, 0.37) + 1000*r(x, 0.41) - 500*r(x, 0.43) +
                       7.5*r(x, 0.5) - 15*r(x, 0.7) + 7.5*r(x, 0.9))}
mu=f(1:Tt/Tt)
mu=mu
beta2=5*mu/sqrt(norm(mu,'2'))

f=function(x){return(0.5 + 2*cos(4*pi*x) + 2*cos(24*pi*x))}
mu=f(1:Tt/Tt)
mu=mu-min(mu)
mu=5*mu/sqrt(norm(mu,'2'))

set.seed(12345)
N=100
Y=matrix(nrow = N,ncol = Tt)
X=matrix(rnorm(N*2),nrow=N,ncol=2)
sigma=0.3
for (i in 1:Tt) {
  Y[,i]=mu[i]+X%*%rbind(beta1[i],beta2[i])+rnorm(N,0,sigma)
}
for (i in 1:N) {
  Y[i,]=rpois(Tt,exp(Y[i,]))
}

r1=wavelet_fda(Y,X,0.1,trans='log')
plot(r1[1,],type='l',ylab='',main='Estimate of Beta')
lines(mu,col='grey80')
plot(r1[2,],type='l',ylab='',main='Estimate of Beta')
lines(beta1,col='grey80')
plot(r1[3,],type='l',ylab='',main='Estimate of Beta')
lines(beta2,col='grey80')

```

```{r,eval=F}

set.seed(12345)
N=100
Y=matrix(nrow = N,ncol = Tt)
X=matrix(rnorm(N*2),nrow=N,ncol=2)
sigma=0.1
for (i in 1:Tt) {
  Y[,i]=mu[i]+X%*%rbind(beta1[i],beta2[i])+rnorm(N,0,sigma)
}
for (i in 1:N) {
  Y[i,]=rpois(Tt,(Y[i,])^2)
}

r1=wavelet_fda(Y,X,0.1,trans='vst')
plot(r1[1,],type='l',ylab='',main='Estimate of Beta')
lines(mu,col='grey80')
plot(r1[2,],type='l',ylab='',main='Estimate of Beta')
lines(beta1,col='grey80')
plot(r1[3,],type='l',ylab='',main='Estimate of Beta')
lines(beta2,col='grey80')
```
