---
title: "Generalized smoothing by adaptive shrinkage"
author: "Dongyue Xie"
date: "2018-10-11"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Poisson likelihood aprroximation

Examples from Matthew. (<http://stephens999.github.io/poisson/analysis/poisson_loglik.html>)

```{r}
plot_poisson_loglik=function(x,range1=1,range2=1,...){
  m = ifelse(x==0,log(0.01),log(x))
  sd_approx = ifelse(x==0,-m/sqrt(2),exp(m)^-0.5)
  l = m + seq(-range1,range2,length=100)
  loglik = dpois(x,exp(l),log=TRUE)
  plot(l, loglik-max(loglik),type="l",...)
  abline(v=log(x))
  normlik = log(dnorm(l,m,sd=sd_approx))
  lines(l,normlik-max(normlik),col="blue")
}
par(mfrow=c(2,2))
plot_poisson_loglik(1,main="x=1")
plot_poisson_loglik(2,main="x=2")
plot_poisson_loglik(20,main="x=20")
plot_poisson_loglik(0,main="x=0",range1=1,range2=5)
```


## Examples of smashgen-Poisson

Mean function range [0.1,6].
```{r,warning=FALSE,message=FALSE}
library(smashrgen)
seed=12345
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 512
t = 1:n/n
m = spike.f(t)

m=m*2+0.1
#range(m)

sig=0

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)

sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))
#################

sig=0.1

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)


sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0.1')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))


#############
sig=0.5

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0.5')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))
########

sig=1

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=1')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))



```

## Examples of smashgen-Poisson

Mean function range [30,90].

```{r,warning=FALSE,message=FALSE}
library(smashrgen)
seed=12345
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 512
t = 1:n/n
m = spike.f(t)

m=m*20+30
#range(m)

sig=0

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)

sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))
#################

sig=0.1

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)


sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0.1')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))

#############
sig=0.3

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)


sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0.3')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))
#############
sig=0.5

set.seed(seed)
lambda=exp(log(m)+rnorm(n,0,sig))
x=rpois(n,lambda)



sgen=smash_gen_lite(x,wave_family = 'DaubLeAsymm',filter.number = 8)
sp=smash.poiss(x)

#par(mfrow=c(2,2))

plot(x,col='grey80',ylab='',xlab='',main='nugget=0.5')
lines(m,col='grey10')
lines(sp,col=2)
lines(sgen,col=4)

legend('topleft',c('data','true mean','smash.pois','smashgen'),lty=c(0,1,1,1),pch=c(1,NA,NA,NA),col=c('grey80','grey10',2,4))
########


```


## Examples of smashgen-Binomial

$n_t=10$, p has range [0.011,0.99]

```{r,warning=FALSE,message=FALSE}
seed=12345
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 512
t = 1:n/n
m = spike.f(t)

p = spike.f(t)*3-4.5

ntri=rep(10,n)

sig=0

set.seed(seed)

ng=rnorm(n,0,sig)
m=exp(p+ng)
q=m/(1+m)
x=rbinom(n,ntri,q)


sgen=smash_gen_lite(x,ntri=ntri,dist_family = 'binomial',wave_family = 'DaubLeAsymm',filter.number = 8)


par(mfrow=c(2,2))

plot(x/ntri,col='grey80',ylab='',xlab='',main='nugget=0')
lines(exp(p)/(1+exp(p)),col='grey10')
lines(sgen,col=4)

legend('topleft',c('data','true mean','smashgen'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey10',4))
#################

sig=0.1

set.seed(seed)

ng=rnorm(n,0,sig)
m=exp(p+ng)
q=m/(1+m)
x=rbinom(n,ntri,q)


sgen=smash_gen_lite(x,ntri=ntri,dist_family = 'binomial',wave_family = 'DaubLeAsymm',filter.number = 8)


#par(mfrow=c(2,2))

plot(x/ntri,col='grey80',ylab='',xlab='',main='nugget=0.1')
lines(exp(p)/(1+exp(p)),col='grey10')
lines(sgen,col=4)

legend('topleft',c('data','true mean','smashgen'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey10',4))

#############
sig=0.5

set.seed(seed)

ng=rnorm(n,0,sig)
m=exp(p+ng)
q=m/(1+m)
x=rbinom(n,ntri,q)


sgen=smash_gen_lite(x,ntri=ntri,dist_family = 'binomial',wave_family = 'DaubLeAsymm',filter.number = 8)


#par(mfrow=c(2,2))

plot(x/ntri,col='grey80',ylab='',xlab='',main='nugget=0.5')
lines(exp(p)/(1+exp(p)),col='grey10')
lines(sgen,col=4)

legend('topleft',c('data','true mean','smashgen'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey10',4))

########

sig=1

set.seed(seed)

ng=rnorm(n,0,sig)
m=exp(p+ng)
q=m/(1+m)
x=rbinom(n,ntri,q)


sgen=smash_gen_lite(x,ntri=ntri,dist_family = 'binomial',wave_family = 'DaubLeAsymm',filter.number = 8)


#par(mfrow=c(2,2))

plot(x/ntri,col='grey80',ylab='',xlab='',main='nugget=1')
lines(exp(p)/(1+exp(p)),col='grey10')
lines(sgen,col=4)

legend('topleft',c('data','true mean','smashgen'),lty=c(0,1,1),pch=c(1,NA,NA),col=c('grey80','grey10',4))


```

## Chip-exo/Chip-seq examples

Data from Rhee and Pugh 2011 and Cuddapah et al. 2009, processed by Kevin.

```{r}
extract_counts_CTCF <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")

  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}
par(mfrow=c(2,2))
##########
# chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipexo_examples/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt")
# 
# smash.out=smash.poiss(chipexo1$counts)
# y=reflect(chipexo1$counts,'both',c(300,299))
# smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)
# 
# plot(chipexo1$counts, col = "gray80", ylab = "rep1 forward", xlab = "", main = "EncodeBroadHistoneHelas - Rep 1")
# lines(smash.out, col = 2, lwd = 1)
# lines(smashgen.out[301:725],col=4,lwd=2)
# 
# legend("topright", # places a legend at the appropriate place
#        c("truth","smash-poiss",'smash-gen'), # puts text in the legend
#        lty=c(0,1,1), # gives the legend appropriate symbols (lines)
#        pch=c(1,NA,NA),
#        lwd=c(1,1,2),
#        cex = 0.5,
#        col=c("gray80","red", "blue"))

#############
chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt")

smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)

plot(chipexo1$counts, col = "gray80", ylab = "rep1 forward", xlab = "", main = "EncodeOpenChromChipHelas - Rep 1")
lines(smash.out, col = 2, lwd = 1)
lines(smashgen.out[301:725],col=4,lwd=2)

legend("topright", # places a legend at the appropriate place
       c("truth","smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(0,1,1), # gives the legend appropriate symbols (lines)
       pch=c(1,NA,NA),
       lwd=c(1,1,2),
       cex = 0.5,
       col=c("gray80","red", "blue"))

###########
chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipexo_examples/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt")

smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)

plot(chipexo1$counts, col = "gray80", ylab = "rep2 forward", xlab = "", main = "EncodeOpenChromChipHelas - Rep 2
")
lines(smash.out, col = 2, lwd = 1)
lines(smashgen.out[301:725],col=4,lwd=2)

legend("topright", # places a legend at the appropriate place
       c("truth","smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(0,1,1), # gives the legend appropriate symbols (lines)
       pch=c(1,NA,NA),
       lwd=c(1,1,2),
       cex = 0.5,
       col=c("gray80","red", "blue"))


##Not very good exampels above ones
#############
chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipseq_examples/example_CTCF_raw_rep1_Forward_counts.txt")

smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)

plot(chipexo1$counts, col = "gray80", ylab = "forward strand", xlab = "", main = "CTCF raw rep 1")
lines(smash.out, col = 2, lwd = 1)
lines(smashgen.out[301:725],col=4,lwd=2)

legend("topright", # places a legend at the appropriate place
       c("truth","smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(0,1,1), # gives the legend appropriate symbols (lines)
       pch=c(1,NA,NA),
       lwd=c(1,1,2),
       cex = 0.5,
       col=c("gray80","red", "blue"))

###############
chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipseq_examples/example_CTCF_raw_rep2_Forward_counts.txt")

smash.out=smash.poiss(chipexo1$counts)
y=reflect(chipexo1$counts,'both',c(300,299))
smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)

plot(chipexo1$counts, col = "gray80", ylab = "forward strand", xlab = "", main = "CTCF raw rep 2")
lines(smash.out, col = 2, lwd = 1)
lines(smashgen.out[301:725],col=4,lwd=2)

legend("topright", # places a legend at the appropriate place
       c("truth","smash-poiss",'smash-gen'), # puts text in the legend
       lty=c(0,1,1), # gives the legend appropriate symbols (lines)
       pch=c(1,NA,NA),
       lwd=c(1,1,2),
       cex = 0.5,
       col=c("gray80","red", "blue"))

###############
# chipexo1 <- extract_counts_CTCF("/Users/dongyue/Documents/smash-gen/data/chipseq_examples/example_CTCF_raw_rep3_Forward_counts.txt")
# 
# smash.out=smash.poiss(chipexo1$counts)
# y=reflect(chipexo1$counts,'both',c(300,299))
# smashgen.out=smash_gen_lite(y,wave_family = 'DaubLeAsymm',filter.number = 8)
# 
# plot(chipexo1$counts, col = "gray80", ylab = "forward strand", xlab = "", main = "CTCF raw rep 3")
# lines(smash.out, col = 2, lwd = 1)
# lines(smashgen.out[301:725],col=4,lwd=2)
# 
# legend("topright", # places a legend at the appropriate place
#        c("truth","smash-poiss",'smash-gen'), # puts text in the legend
#        lty=c(0,1,1), # gives the legend appropriate symbols (lines)
#        pch=c(1,NA,NA),
#        lwd=c(1,1,2),
#        cex = 0.5,
#        col=c("gray80","red", "blue"))
```

## More details

Please click <https://dongyuexie.github.io/smash-gen/> for more details.

## Reference

Stephens, M. (2017). False discovery rates: a new deal. Biostatistics, 18(2), 275-294.

Xing, Z., & Stephens, M. (2016). Smoothing via Adaptive Shrinkage (smash): denoising Poisson and heteroskedastic Gaussian signals. arXiv preprint arXiv:1605.07787.

Johnstone, I. M., & Silverman, B. W. (2005). Empirical Bayes selection of wavelet thresholds. Annals of Statistics, 1700-1752.

Rhee, H. S., & Pugh, B. F. (2011). Comprehensive genome-wide protein-DNA interactions detected at single-nucleotide resolution. Cell, 147(6), 1408-1419.

Cuddapah, S., Jothi, R., Schones, D. E., Roh, T.-Y., Cui, K., & Zhao, K. (2009). Global analysis of the insulator binding protein CTCF in chromatin barrier regions reveals demarcation of active and repressive domains. Genome Research, 19(1), 24–32. http://doi.org/10.1101/gr.082800.108
