---
title: "Variance stablizing transformation"
author: "Dongyue Xie"
date: "2018-10-13"
output: 
  workflowr::wflow_html:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F,warning = F)
```

## Introduction

Variance stablizing transformation.

$E(X)=\mu$ and $Var(X)=g(\mu)$, want to find $f(\cdot)$ s.t $Var(f(X))$ has constant variance. Consider the Taylor series expansion of $f(X)$ around $\mu$: $f(X)\approx f(\mu)+(Y-\mu)f'(\mu)$ so we have $[f(X)-f(\mu)]^2\approx (X-\mu)^2(f'(\mu))^2 \Rightarrow Var(f(X))\approx Var(X)(f'(\mu))^2$. 

For poisson distribution, $(f'(\mu))^2\propto \mu^{-1}$ so if we take $Y=\sqrt{X}$ then $Var(Y)\approx \frac{1}{4}$. This was original proposed by Bartlett in 1936.

For Binomial data, $(f'(\mu))^2\propto 1/(np(1-p))$ so if we take $Y=sin^{-1}(\sqrt{X/n})$ then $Var(Y)\approx \frac{1}{2}$.

Anscombe(1948) shows that for $Y=\sqrt{X+c}$, $Var(Y)\approx \frac{1}{4}[1+\frac{3-8c}{8\mu}+\frac{32c^2-52c+17}{2\mu^2}]]$. If take $c=3/8$ and for large $\mu$, $Var(Y)\approx 1/4$. Also clearly, $\lim_{\mu\to 0}Var(\sqrt{X+c})=0$. 

```{r}
mu=c(seq(0,1,length.out = 50),seq(1,10,length.out = 50))
ans=c()
sr=c()
set.seed(12345)
for (i  in 1:100) {
  x=rpois(1e6,mu[i])
  ans[i]=var(sqrt(x+3/8))
  sr[i]=var(sqrt(x))
}
plot(mu,ans,type='l',ylim=c(0,0.5),ylab='')
lines(mu,sr,col=4)
abline(a=0.25,b=0,lty=2)
legend('bottomright',c('anscombe','square root'),lty=c(1,1),col=c(1,4))
```

# Compare log and anscombe transformation

Poisson variance stablizing trasformations: square root and Anscombe transformation.

For vst, if we observe $x=0$, then I use $var(\sqrt{X+3/8})=0$ instead of $1/4$.

```{r}
vst_smooth=function(x,method,ep=1e-5){
  n=length(x)
  if(method=='sr'){
    x.t=sqrt(x)
    x.var=rep(1/4,n)
    x.var[x==0]=0
    mu.hat=(smashr::smash.gaus(x.t,sigma=sqrt(x.var)))^2
    
  }
  if(method=='anscombe'){
    x.t=sqrt(x+3/8)
    x.var=rep(1/4,n)
    x.var[x==0]=0
    mu.hat=(smashr::smash.gaus(x.t,sigma=sqrt(x.var)))^2-3/8
  }
  if(method=='log'){
    x.t=x
    x.t[x==0]=ep
    x.var=1/x.t
    x.t=log(x.t)
    mu.hat=exp(smashr::smash.gaus(x.t,sigma=sqrt(x.var)))
  }
  return(mu.hat)
}


```

```{r}
simu_study=function(m,sig=0,nsimu=30,seed=12345){
  set.seed(seed)
  sr=c()
  an=c()
  ashp=c()
  for (i in 1:nsimu) {
    lambda=exp(log(m)+rnorm(n,0,sig))
    x=rpois(n,lambda)
    mu.sr=vst_smooth(x,'sr')
    mu.an=vst_smooth(x,'anscombe')
    mu.ash=smash_gen_lite(x,sigma = sig)
    sr=rbind(sr,mu.sr)
    an=rbind(an,mu.an)
    ashp=rbind(ashp,mu.ash)
  }
  return(list(sr=sr,an=an,ashp=ashp))
}

```

When there are a number of $0s$ in the observation: (mean function (0.1,6))

```{r,fig.width=12,fig.height=8}
library(ashr)
library(smashrgen)
spike.f = function(x) (0.75 * exp(-500 * (x - 0.23)^2) + 1.5 * exp(-2000 * (x - 0.33)^2) + 3 * exp(-8000 * (x - 0.47)^2) + 2.25 * exp(-16000 * 
    (x - 0.69)^2) + 0.5 * exp(-32000 * (x - 0.83)^2))
n = 512
t = 1:n/n
m = spike.f(t)

m=m*2+0.1
range(m)


result=simu_study(m)


mses=lapply(result, function(x){apply(x, 1, function(y){mean((y-m)^2)})})

par(mfrow=c(2,2))

for (j in 1:4) {
  plot(m,type='l',main='nugget=0')
lines(result$sr[j,],col=2)
lines(result$an[j,],col=3)
lines(result$ashp[j,],col=4)
legend('topright',c('mean','square root','anscombe','smashgen'),lty=c(1,1,1,1),col=c(1,2,3,4))

}
```

```{r}
boxplot(mses,names = c('square root','anscombe','smashgen'),main='nugget=0',ylab='MSE')
```

```{r,fig.width=12,fig.height=8}
########
result=simu_study(m,sig=1)


mses=lapply(result, function(x){apply(x, 1, function(y){mean((y-m)^2)})})

#unlist(lapply(mses, mean))

par(mfrow=c(2,2))
for (j in 1:4) {
  plot(m,type='l',main='nugget=1')
#lines(result$sr[1,],col=2)
lines(result$an[j,],col=3)
lines(result$ashp[j,],col=4)
legend('topright',c('mean','anscombe','smashgen'),lty=c(1,1,1),col=c(1,3,4))
#legend('topright',c('mean','square root','anscombe','smashgen'),lty=c(1,1,1,1),col=c(1,2,3,4))

}

```

```{r}
boxplot(mses,names = c('square root','anscombe','smashgen'),main='nugget=1',ylab='MSE')

###############
```

\text{Increase range of mean function to (30,150)}

```{r,fig.width=12,fig.height=8}

m=m*20+30
range(m)

result=simu_study(m)

mses=lapply(result, function(x){apply(x, 1, function(y){mean((y-m)^2)})})

#unlist(lapply(mses, mean))

par(mfrow=c(2,2))

for (j in 1:4) {
  plot(m,type='l',main='nugget=0')
lines(result$sr[j,],col=2)
lines(result$an[j,],col=3)
lines(result$ashp[j,],col=4)
legend('topright',c('mean','square root','anscombe','smashgen'),lty=c(1,1,1,1),col=c(1,2,3,4))

}

```

```{r}
boxplot(mses,names = c('square root','anscombe','smashgen'),main='nugget=0',ylab='MSE')
```

## Step function

mean: (1,7)

```{r,fig.width=12,fig.height=8}
m=c(rep(1,n/4),rep(5,n/4),rep(7,n/4),rep(1,n/4))

result=simu_study(m)

mses=lapply(result, function(x){apply(x, 1, function(y){mean((y-m)^2)})})

#unlist(lapply(mses, mean))

par(mfrow=c(2,2))

for (j in 1:4) {
  plot(m,type='l',main='nugget=1',ylim=c(-1,8))
#lines(result$sr[1,],col=2)
lines(result$an[j,],col=3)
lines(result$ashp[j,],col=4)
legend('topright',c('mean','anscombe','smashgen'),lty=c(1,1,1),col=c(1,3,4))
#legend('topright',c('mean','square root','anscombe','smashgen'),lty=c(1,1,1,1),col=c(1,2,3,4))

}
```

```{r}
boxplot(mses,names = c('square root','anscombe','smashgen'),main='nugget=0',ylab='MSE')

```

## Heavi Sine function

mean: (1,24)

```{r,fig.width=12,fig.height=8}
m=wavethresh::DJ.EX(n)
m=m$heavi+15
range(m)

result=simu_study(m)

mses=lapply(result, function(x){apply(x, 1, function(y){mean((y-m)^2)})})

#unlist(lapply(mses, mean))

par(mfrow=c(2,2))

for (j in 1:4) {
  plot(m,type='l',main='nugget=1')
#lines(result$sr[1,],col=2)
lines(result$an[j,],col=3)
lines(result$ashp[j,],col=4)
legend('topright',c('mean','anscombe','smashgen'),lty=c(1,1,1),col=c(1,3,4))
#legend('topright',c('mean','square root','anscombe','smashgen'),lty=c(1,1,1,1),col=c(1,2,3,4))

}
```

```{r}
boxplot(mses,names = c('square root','anscombe','smashgen'),main='nugget=0',ylab='MSE')

```
