---
title: "Iterative vst"
author: "Dongyue Xie"
date: "2018-10-26"
output: workflowr::wflow_html

---

## Smoothing exp family data

Observations are $x$ count data which are modeled using Poisson distribution. Assume $X\sim Pois(\mu)$, and $\mu$ is smooth or 'noisy' smooth. Noisy smooth means $\mu$ is conraminated by noise. We model it as $\mu=h(m+\epsilon)$, where $\epsilon$ is a random error $\epsilon\sim N(0,\sigma^2)$ and $h(\cdot)$ is a one-to-one continuous real valued function. The question is how to recover the smooth structure in $\mu$.

Variance stabilizing transformation approach:

Derivation of anscombe transformation:

suppose $X\sim Pois(\mu)$ and define $t=X-\mu$, $m=\mu+c$. Expand $Y=\sqrt{X+c}=\sqrt{m+t}$ around $t=0$ we have $Y=m^{1/2}+\frac{1}{2}m^{-1/2}t-\frac{1}{8}m^{-3/2}t^2+\frac{1}{16}m^{-5/2}t^3+O(t^4)$.

Moments of $t$: $E(t)=0$, $E(t^2)=\mu$, $E(t^3)=\mu$, $E(t^4)=3\mu^2+\mu$, $E(t^6)=-\mu^4+15\mu^3+25\mu^2+\mu$

Then the variance of $Y$ is(using up to $t^2$)  $Var(Y)=\frac{1}{4}m^{-1}\mu+\frac{1}{64}m^{-3}(2\mu^2+\mu)-\frac{1}{8}m^{-2}\mu$ and if $\mu\to\infty$, $Var(Y)\to 1/4$.

**How did Anscombe make $c$ in nominator???** I think is to use higher order of t and let $\mu\to\infty$.

$Y=\sqrt{X+c}$ where $c$ is a constant, $c\geq 0$ and then $Y\sim N(\sqrt{\mu+c},1/4)$ for large $\mu$. ($\mu\geq 4$ is already very good approx; $\mu\geq 2$ is ok.) We can write $Y=\sqrt{\mu+c}+N(0,1/4)$ and so $Y=m+\epsilon+N(0,1/4)$. If $\sigma^2$ is known, we can apply any Gaussian non-parametric smoothing methods to estiamte $m$; if unkown, we might want to estiamte $\sigma^2$ or $\sigma^2+1/4$ first.

One problem is that $Var(Y)\approx 1/4$ only holds for large enough $\mu$ and what if we have observations like $x=0$. One immidiate strategy is to use 0 variance for 0 $x$. But the probability of observing $x=0$ when $\mu=1,2$ is 0.37, 0.14 respectively. So we may ask if there are better way to do this.
```{r}
dpois(0,1)
dpois(0,2)
dpois(0,3)
```

One way is to approximate the variance using $\mu$. Then we need a formula bwt var and mean. From the plot below, the simulated variance line is from $10^5$ ramdom samples so it can be regared as the 'true' variance of $Y=\sqrt{X+3/8}$; formula in anscombe's paper is obtained when $\mu\to\infty$ so it cannot deal with small mean; using 2nd order taylor series expnasion always under-estimate the variance. So we might need higher order approximation but it's much more complicated.

```{r}
# mu=1
c=3/8
# m=mu+c
# t=seq(-m,m,length.out = 1000)
# plot(t,sqrt(t+m),type='l')
# #taylor series expnasion around t=0
# taylor=sqrt(m)+1/sqrt(m)*t/2-m^(-1.5)/8*t^2+m^(-2.5)/16*t^3
# lines(t,taylor,col=4)

mu=seq(0,10,length.out = 1000)
m=mu+c
var.y=mu/(4*m)+(2*mu^2+mu)/(64*m^3)-mu/(8*m^2)
var.ans=1/4*(1+(32*c^2-52*c+17)/(32*mu^2))

var.sim=c()
set.seed(12345)
for (i  in 1:length(mu)) {
  x=rpois(1e6,mu[i])
  var.sim[i]=var(sqrt(x+3/8))
}

plot(mu,var.y,type='l',ylim=c(0,0.5))
lines(mu,var.sim,col=3)
lines(mu,var.ans,col=4)
abline(a=0.25,b=0,lty=2)
legend('bottomright',c('anscombe formula','2nd order taylor','simulated'),col=c(4,1,3),lty=c(1,1,1))
```


## Functional mixed models

A good [tutorial](https://www.mailman.columbia.edu/sites/default/files/pdf/fda_shortcourse.pdf) on functional data analysis.

shim&stephens:

This is a 'Function on scalar regression' case: response(count) is functional data and covariate(genotype) is a scalar. For each subject $i=1,...,n$, $y_i(s)=x_i^T\beta(s)+\epsilon_i(s)$, where $s$ is the grid of time points, $|s|=T$ the total length of observations of subject $i$. For example, if $x_i=(1,x_{i1})$, then $y_i(s)=\beta_0(s)+x_{i1}\beta_1(s)+\epsilon_i(s)$. So $\beta_0(s)$ is global mean function of all subjects and $\beta_1(s)$ is global coefficients function of all subjects. We can write the model as $Y=XB+E$ where $T\in R^{n*T}$, $X\in R^{n*p}$, $B\in R^{p*T}$ and $E$ is n by T error matrix.

So the idea for dealing with data in shim and stephens, taking transformation and nugget effect into consideration is:

suppose we observe a sequence of $x$ from an exponential family distribution, we do a transformation on X and obtain $y=h(x)=m+\beta z+\epsilon_1+\epsilon_2$ where $m$ has a smooth structure, $\epsilon$s are normal distributed random errors. Then apply DWT to $y$ yeild $Wy=Wm+W\beta z+W\epsilon_1+W\epsilon_2$. Then we can use similar methods in paper to do inference. A problem is how to estimate nugget effect.
